Pre-/Post-Trikafta in CFRD

### 2022-10-06 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
Update pre_post_trikafta_analysis.Rmd

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lubridate)
source("~/GitHub/CF/get_genotype_class_severity.R")
variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
              "G-tube in past 12 months","Start","Date","Age")
```

```{r include=FALSE}
# CHCO
chco_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv",show_col_types = F)
chco_a1c$Date = parse_date_time(chco_a1c$Date,orders = "mdy")
chco_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv",show_col_types = F)
chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
chco = full_join(chco_a1c,chco_bmi,by = c("MRN", "Date"))
# Fill CFF ID by MRN
chco= chco %>% group_by(MRN) %>% 
  fill(`CFF ID`,.direction = "downup") %>% ungroup() %>%
  filter(!is.na(`CFF ID`)) %>%
  select(-MRN)
# Add demographics
chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
chco_reg = chco_reg %>% select(-MRN)
chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
chco = left_join(chco,chco_reg,by = "CFF ID")
# Add Time
chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
```

```{r include=FALSE}
# Montana
mon_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv",show_col_types = F)
mon_bmi$Date = parse_date_time(mon_bmi$Date,orders = "mdy")
mon_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv",show_col_types = F)
mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
# Get start date and CFRD status
mon_starts = mon_a1c %>% select(`CFF ID`,`CFRD Status`,Start) %>% 
  distinct() %>% filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia")
mon_starts$Start = parse_date_time(mon_starts$Start,orders = "mdy")
# Remove unnecessary columns
mon_a1c = mon_a1c %>% select(`CFF ID`,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
# Merge and remove those without CFRD or start date
mon = full_join(mon_bmi,mon_a1c,by = c("CFF ID", "Date"))
mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start), by = "CFF ID")
# Add Time
mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
```

```{r include=FALSE}
# UM
um_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",show_col_types = F)
um_a1c$Age = round(um_a1c$Age / 365.25,1)
um_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",show_col_types = F)
um_bmi$Age = round(um_bmi$Age / 365.25,1)
um_ogtt = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",show_col_types = F)
um_ogtt$Age = round(um_ogtt$Age / 365.25,1)
um_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",show_col_types = F)
um_pft$Age = round(um_pft$Age / 365.25,1)
um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
# Join
um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
um = left_join(um,um_race,by = "CFF ID")
```

```{r include=FALSE}
# UW
uw_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv",show_col_types = F)
uw_bmi$Date = parse_date_time(uw_bmi$Date,orders = "mdy")
uw_glyc = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv",show_col_types = F)
uw_glyc$Date = parse_date_time(uw_glyc$Date,orders = "mdy")
uw_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv",show_col_types = F)
uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
uw_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_registry.csv",show_col_types = F)
uw_reg$Start = parse_date_time(uw_reg$Start,orders = "mdy")
# Merge
uw = reduce(list(uw_bmi,uw_glyc,uw_pft,uw_reg),full_join)
```

```{r include=FALSE}
# Put everything together

```

# Participant Characteristics

```{r results='asis'}
t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
summary(t1)
```

>>>>

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(lubridate)
library(rspiro)
source("~/GitHub/CF/get_genotype_class_severity.R")
t1_variables = c("CFF ID","Site","Sex","Race","Genotypes1","Genotypes2","Pancreatic Status",
              "G-tube in past 12 months","Time","Age","BMI","Height","Weight","FEV1","FVC")
```

```{r include=FALSE}
# CHCO
chco_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_glycemic.csv",show_col_types = F)
chco_a1c$Date = parse_date_time(chco_a1c$Date,orders = "mdy")
chco_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv",show_col_types = F)
chco_bmi$Date = parse_date_time(chco_bmi$Date,orders = "mdy")
chco = full_join(chco_a1c,chco_bmi)
# Fill CFF ID by MRN
chco= chco %>% group_by(MRN) %>% 
  fill(`CFF ID`,.direction = "downup") %>% ungroup() %>%
  filter(!is.na(`CFF ID`)) %>%
  select(-MRN)
# Add demographics
chco_reg = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_registry.csv",show_col_types = F)
chco_reg$Start = parse_date_time(chco_reg$Start,orders = "mdy")
chco_reg$DOB = parse_date_time(chco_reg$DOB,orders = "mdy")
keep = chco_reg %>% filter(!is.na(`CFRD Diagnosis Date`)) %>% pull(`CFF ID`)
exclude = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/chco_insulin.csv",show_col_types = F)
exclude = exclude %>% filter(!is.na(`Remove (see notes)`)) %>% pull(MRN)
exclude = chco_reg$`CFF ID`[match(exclude,chco_reg$MRN)]
keep =setdiff(keep,exclude)
chco = left_join(chco,chco_reg) %>% filter(`CFF ID` %in% keep)
# Add missing variables
chco$Age = round(as.numeric(difftime(chco$Date,chco$DOB,units = "days"))/365.25,1)
chco$Time = as.numeric(difftime(chco$Date,chco$Start,units = "days"))
chco$Site = "CHCO"
```

```{r include=FALSE}
# Montana
mon_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv",show_col_types = F)
mon_bmi$Date = parse_date_time(mon_bmi$Date,orders = "mdy")
mon_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv",show_col_types = F)
mon_a1c$Date = parse_date_time(mon_a1c$Date,orders = "mdy")
mon_a1c$DOB = parse_date_time(mon_a1c$DOB,orders = "mdy")
# Get start date and CFRD status
mon_starts = mon_a1c %>% select(`CFF ID`,`CFRD Status`,Start) %>% 
  distinct() %>% filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia")
mon_starts$Start = parse_date_time(mon_starts$Start,orders = "mdy")
# Remove unnecessary columns
mon_a1c = mon_a1c %>% select(`CFF ID`,DOB,Date,HbA1c,`OGTT Fasting`,`OGTT Two Hour`)
# Merge and remove those without CFRD or start date
mon = full_join(mon_bmi,mon_a1c)
mon= mon %>% filter(`CFF ID` %in% mon_starts$`CFF ID`)
mon = left_join(mon,mon_starts %>% select(`CFF ID`,Start))
# Add demographics
mon_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_demo.csv",show_col_types = F)
mon_demo = mon_demo %>% select(`CFF ID`,Race,Sex)
mon = left_join(mon,mon_demo)
# Add genotypes
mon_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_genotypes.csv",show_col_types = F)
mon = left_join(mon,mon_geno)
# Add pancreatic and g tube status
mon_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_pancreatic_insufficient.csv",show_col_types = F)
mon_pan = mon_pan %>% select(-Date) %>% distinct()
mon_pan$`Pancreatic Status` = "Insufficient"
mon_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/montana_gtubes.csv",show_col_types = F)
mon_gtube = mon_gtube %>% select(-Feeding,-Date) %>% distinct()
mon_gtube$`G-tube in past 12 months` = "Yes"
mon = left_join(mon,mon_pan)
mon = left_join(mon,mon_gtube)
# Add missing variables
mon = mon %>% group_by(`CFF ID`) %>% fill(Start,DOB,.direction = "downup")
mon$Time = as.numeric(difftime(mon$Date,mon$Start,units = "days"))
mon$Age = round(as.numeric(difftime(mon$Date,mon$DOB))/365.25,1)
mon$Site = "Montana"
```

```{r include=FALSE}
# UM
um_a1c = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",show_col_types = F)
um_a1c$Age = round(um_a1c$Age / 365.25,1)
um_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",show_col_types = F)
um_bmi$Age = round(um_bmi$Age / 365.25,1)
um_ogtt = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",show_col_types = F)
um_ogtt$Age = round(um_ogtt$Age / 365.25,1)
um_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",show_col_types = F)
um_pft$Age = round(um_pft$Age / 365.25,1)
um_race = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_race.csv",show_col_types = F)
um_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",show_col_types = F)
keep = um_demo %>% 
  filter(`CFRD Status` =="CFRD with or without fasting hyperglycemia") %>%
  pull(`CFF ID`)
um_demo = um_demo %>% select(-`CFRD Status`) %>% 
  rename(`G-tube in past 12 months`=`G tube in past 12 months`)
# Join
um = reduce(list(um_a1c,um_bmi,um_ogtt,um_pft),full_join)
um = left_join(um,um_race)
um = left_join(um,um_demo)
um = um %>% filter(`CFF ID` %in% keep) %>% rename(Site = "site")
```

```{r include=FALSE}
# UW
uw_bmi = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_bmi.csv",show_col_types = F)
uw_bmi$Date = parse_date_time(uw_bmi$Date,orders = "mdy")
uw_glyc = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_glycemic.csv",show_col_types = F)
uw_glyc$Date = parse_date_time(uw_glyc$Date,orders = "mdy")
uw_pft = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pft.csv",show_col_types = F)
uw_pft$Date = parse_date_time(uw_pft$Date,orders = "mdy")
uw_demo = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_demo.csv",show_col_types = F)
uw_demo = uw_demo %>% select(`CFF ID`,Race,Sex)
# Get CFRD status and start date
uw_start = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_mods.csv",show_col_types = F)
uw_start$`Date of Encounter` = parse_date_time(uw_start$`Date of Encounter`,orders = "mdy")
uw_start = uw_start %>% arrange(`CFF ID`,`Date of Encounter`) %>% 
  group_by(`CFF ID`) %>% filter(row_number()==n()) %>%
  select(`CFF ID`,`Date of Encounter`) %>% rename(Start = `Date of Encounter`)
uw_cfrd = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_cfrd.csv",show_col_types = F)
uw_cfrd = uw_cfrd %>% 
  filter(`CFRD Status`=="CFRD with or without fasting hyperglycemia") %>% 
  select(`CFF ID`) %>% distinct() %>% pull(`CFF ID`)
keep = unique(uw_cfrd,uw_start$`CFF ID`)
# Add genotypes
uw_geno = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_genotypes.csv",show_col_types = F)
# Pancreatic status
uw_pan = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv",show_col_types = F)
uw_pan = uw_pan %>% select(-Date) %>% distinct() %>% mutate(`Pancreatic Status`="Insufficient")
# G tubes
uw_gtube = read_csv("/Volumes/Documents/Work/Vigers/CF/Colleen Wood/Prepost Triakfta/Data_Cleaned/uw_gtubes.csv",show_col_types = F)
uw_gtube = uw_gtube %>% select(`CFF ID`) %>% distinct() %>% mutate(`G-tube in past 12 months`="Yes")
# Merge
uw = reduce(list(uw_bmi,uw_glyc,uw_pft),full_join)
uw = left_join(uw,uw_demo)
uw = left_join(uw,uw_geno)
uw = left_join(uw,uw_pan)
uw = left_join(uw,uw_gtube)
uw = left_join(uw,uw_start) %>% filter(`CFF ID` %in% keep)
# Other variables (no weight for this site)
uw$Time = as.numeric(difftime(uw$Date,uw$Start,units = "days"))
uw$Site = "UW"
uw$Weight = NA
```

```{r include=FALSE}
# Put everything together
chco = chco %>% select(all_of(t1_variables))
mon = mon %>% select(all_of(t1_variables))
um = um %>% select(all_of(t1_variables))
uw = uw %>% select(all_of(t1_variables))
df = bind_rows(chco,mon,um,uw) %>% arrange(`CFF ID`,Time)
# Age group
df$`Age Group` = cut(df$Age,breaks = c(-Inf,20,Inf),labels = c("Pedatric","Adult"),right = F)
# Race and sex
df$Race = factor(df$Race,levels = c("Black or African American","Caucasian",
                                    "Other","Some other race","Two or more races","White"),
                 labels = c("Black","Caucasian",
                                    "Other","Other","Other","Caucasian"))
df$Sex = factor(df$Sex,levels = c("Male","Female"))
# BMI and FEV calculations
age = df$Age
height = df$Height/100
race = df$Race
old <- c("Caucasian","Black","Other")
new <- c(1,2,5)
race = new[match(race, old)]
gender = as.numeric(df$Sex)
df$ppFEV1  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FEV1 = df$FEV1)
df$ppFVC  = pctpred_GLI(age,height,gender = gender,ethnicity = race,FVC = df$FVC)
# Genotype severity
df$Genotypes1 = get_cf_genotype_class_severity(df$Genotypes1)
df$Genotypes2 = get_cf_genotype_class_severity(df$Genotypes2)
df = df %>% unite(Genotype,Genotypes1,Genotypes2)
df$Genotype = factor(df$Genotype,levels = c("Mild_Severe","Severe_Mild","Severe_Severe",
                                            "Severe_Unknown","Unknown_Unknown"),
                     labels = c("Low","Low","High","Unknown","Unknown"))
```

# Participant characteristics at date closest to CFTR modulator start

```{r results='asis'}
t1_data = df %>% group_by(`CFF ID`) %>% slice_min(abs(Time))

t1 = tableby(~Age+BMI.Percentile+Weight+Sex+,pre_post[pre_post$Before_After == "Before",])
summary(t1)
```
